Load packages
```{r}
packages = c("corrplot","ggpubr","GGally","tidyverse", "nnet","caret", "MLmetrics", "rpart.plot","ggplot2")
for (p in packages) {
  # 'require' checks whether package exists and returns TRUE or FALSE
  if (!require(p, character.only = T)) {
    install.packages(p)
  }
  library(p, character.only = T)
}
setwd('/Users/runyuwang/Google Drive/Visual Analytics/Assignment/Assignment Workspace/')

all_bird_file = 'all_birds_wav.csv'
test_bird_file = 'test_birds_from_kasios_wav.csv'
```

Load all bird file and test bird file
```{r}
all_bird_data = read_csv(all_bird_file)
test_bird_data = read_csv(test_bird_file) 
```

Add English Name Column for all bird dataframe
```{r}
#find string before .wav
new_cols= str_split_fixed(all_bird_data$sound, '.wav', n = Inf)

#remove file id (numerical value)
new_cols =as.data.frame(word(all_bird_data$sound,1,sep = "\\d"))

#assign column name
colnames(new_cols) = 'English_Name'

#remove delimiter
#convert English Name to char, as below needs to use char to compare
data = data.frame(lapply(new_cols, function(x) {gsub("-", " ", x)}),stringsAsFactors=FALSE)
all_bird_data = cbind(data, all_bird_data)
```

EDA, Remove NA and Remove Column Sound
```{r}
#drop columns with NA value
drops = c('pitchCep_mean', 'pitchCep_median', 'pitchCep_sd')
all_bird_data = all_bird_data[,!names(all_bird_data) %in% drops]
test_bird_data = test_bird_data[,!names(test_bird_data) %in% drops]

#count number of NA values per column in all_bird_data
colSums(is.na(all_bird_data))
#explore NA value in each row
all_bird_data[rowSums(is.na(all_bird_data))>0,]

#count number of NA values per column in test_bird_data
colSums(is.na(test_bird_data))
#explore NA value in each row
test_bird_data[rowSums(is.na(test_bird_data))>0,]

#omit rows with NA value
all_bird_data <- na.omit(all_bird_data)

#remove column sound
drops = c('sound')
all_bird_data = all_bird_data[,!names(all_bird_data) %in% drops]
test_bird_data=test_bird_data[,!names(all_bird_data) %in% drops]
```


```{r}
#create target variable for Pipit or Not Pipit
all_bird_data$target = 'not rose pipit'
all_bird_data$target[all_bird_data$English_Name == 'Rose Crested Blue Pipit '] = 'rose pipit'

#move target column to the 1st position
all_bird_data = all_bird_data[,c(70, 1:69)]

data_x = all_bird_data[,-c(1,2)]
data_y = all_bird_data[1]
```

split data
```{r}
set.seed(2018)
data_y$target = as.factor(data_y$target)

#merge data_y and data_x for data part
train_data= cbind(data_y, data_x)
data_part <- createDataPartition(y = train_data$target, 
                                 p = 0.7, list = F)

x_train = train_data[data_part,][,-1]
y_train = train_data[data_part,][1]

x_test = train_data[-data_part,][,-1]
y_test = train_data[-data_part,][1]
```

classification model 1: random Forest
```{r}
trctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)

rf <- train(x= x_train, y=y_train[,1], method = "rf",
                 trControl=trctrl,
                 tuneLength = 10)

y_pred = predict(rf,x_test)

Accuracy(y_pred, y_test$target)
Precision(y_test$target,y_pred)
Recall(y_test$target,y_pred)
```

classification model 2: decision tree
```{r}
trctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)

dt <- train(x= x_train, y=y_train[,1], method = "rpart",
                 trControl=trctrl,
                 tuneLength = 10)

y_pred = predict(dt,x_test)

Accuracy(y_pred, y_test$target)
Precision(y_test$target,y_pred)
Recall(y_test$target,y_pred)
```
visualize decision tree
```{r}
rpart.plot(dt$finalModel)
```

use random forest to predict test audio files
```{r}
result_rf = predict(rf,test_bird_data[-1])
cbind(test_bird_data[1], result_rf)
```

use decision tree to predict test audio files
```{r}
result_dt = predict(dt,test_bird_data[-1])
cbind(test_bird_data[1], result_dt)
```

Visual Analytics Method
Remove highly correlated columns at first, too many variables makes it diffcult to visualize
```{r}
data_x = all_bird_data[,c(-1,-2)]
data_y = all_bird_data[1]
data_en_name = all_bird_data[2]

corr_mat = cor(data_x)
high_corr = findCorrelation(corr_mat, cutoff = 0.6, exact = FALSE)
data_x_reduced = data_x[,-c(high_corr)]

```

Apply Corrplot
```{r}
y = cor(data_x_reduced)
corrplot(y, order = "hclust", addrect = 12)
```

remove features based on corr plot observations
Corrplot, select variables, we not only consider to check the variables nearby, but all the variables
For example, group of 'specSlape_sd', 'harmonics_median', 'peakFreq_median', I choose 'peakFreq_median'.

```{r}
drops = c('specSlope_sd','harmonics_median', 'f2_width_median', 'pitchSpec_median', 'f1_width_sd', 'peakFreqCut_sd','dom_sd','quartile25_sd','f3_width_sd','HNR_sd', 'pitchSpec_sd', 'harmonics_sd', 'pitch_sd')
data_x_reduced = data_x_reduced[,!names(data_x_reduced) %in% drops]

#drop variables which are not ideal for visual analytics
#reduce number of features to a managable level
drops = c('amplVoiced_median', 'duration', 'entropy_mean','entropy_sd', 'HNR_median','specSlope_median','voiced' )
data_x_reduced = data_x_reduced[,!names(data_x_reduced) %in% drops]
```

trellis density plot
```{r}
data_reduced = cbind(data_y, data_en_name,data_x_reduced)

data_plot = data_reduced %>%
  group_by(English_Name) %>%
  gather(variable, value, colnames(data_x_reduced)) %>% 
  dplyr::select(English_Name,variable,value)

ggplot(data_plot,aes(x=value,y=..scaled..,fill=variable))+
  geom_density()+
  facet_grid(English_Name~variable,scales='free')+
    theme(strip.text.y = element_text(angle = 0, hjust = 0)) +
    theme(legend.position="none")

```

